---
title: "Charts for report and ppt"
output: html_notebook
---



```{r}
library(tidyverse)
library(bigrquery)
library(scales)
```
Chart 1 - Available observations
```{r}
availability_query <- paste0("SELECT ", 
                        paste0("COUNT(", numeric_vars, ") AS ", numeric_vars, collapse = ", "), " ",
                        "FROM `case-study-2-nec-team.logs.raw_view`")

availability <- bq_project_query("case-study-2-nec-team", availability_query) %>% bq_table_download() %>% 
  pivot_longer(., names(.), names_to = "variable", values_to = "count") 


```

```{r}
availability %>% 
  #remove computed variables 
  .[c(1:7, 10:11), ] %>% 
  mutate(variable = fct_reorder(variable, count)) %>% 
  ggplot() + geom_col(aes(x = variable, y = count/1E6)) + coord_flip() + 
  labs(title = "Number of valid observations", 
       subtitle = "(millions)", x = NULL, y = NULL) + 
  theme(title = element_text(size = 8))

ggsave("final_plots/plot1_report.png", width = 8, height = 8, units = "cm")
# ggsave("final_plots/plot1_ppt.png", width = 22, height = 14, units = "cm")
```

Leave out chart 2 with annotations, go straight to the multi-panel one.
```{r}

lct <- Sys.getlocale("LC_TIME"); 
Sys.setlocale("LC_TIME", "C")

p <- time_series %>% filter(variable %in% c("connections", "CL_RTT", "CL_Retrans", "OS_Retrans", "download_speed_CL_clean", "download_speed_OS_clean") & time < "2018-09-07 18:00:00") %>% 
  #change variable to ordered factor to manually define order in chart
  mutate(variable = factor(variable, c("connections", "CL_RTT", "CL_Retrans", "download_speed_CL_clean", "download_speed_OS_clean", "OS_Retrans"))) %>% 
  ggplot() + geom_line(aes(time, value)) + facet_grid(rows = vars(variable), scales = "free_y") +
  labs(title = "Connections and QoS per hour", x = NULL, y = NULL) +
  #make strip texts smaller, remove y scales
  theme(title = element_text(size = 9),
        strip.text.y = element_text(size = 5), 
        axis.text.y = element_blank(), axis.ticks = element_blank())

p + scale_x_datetime(date_breaks = "8 hours", date_labels = "%a\n%H")
ggsave("final_plots/plot2_report.png", width = 8, height = 16, units = "cm")

p +   scale_x_datetime(date_breaks = "4 hours", date_labels = "%a\n%H")
ggsave("final_plots/plot2_ppt.png", width = 22, height = 14, units = "cm")

Sys.setlocale("LC_TIME", lct)
```


```{r}
cdfs_all %>% filter(variable == "bytesReceived") %>% 
  ggplot() + geom_line(aes(value, CDF), size = 2) + 
  scale_x_log10(labels = label_bytes()) + 
  labs(x = NULL, y = "cumulative distribution",
       title = "CDF for bytesReceived",  caption = "Note: values on log scale.") 

ggsave("final_plots/plot3_report.png", width = 8, height = 8, units = "cm")
ggsave("final_plots/plot3_ppt.png", width = 10, height = 10, units = "cm")
```

Qos by load

```{r}
breakdown_load  %>% filter(variable %in% c("download_speed_CL_clean", "OS_RTT")) %>% 
  ggplot() + geom_line(aes(value, CDF, colour = load), size = 2, alpha = 0.6) + scale_x_log10() + 
  scale_colour_manual(values=c("black", "red", "green")) +
  facet_wrap(vars(variable), nrow = 2, scales = "free_x") +
  labs(x = NULL, y = "cumulative distribution", 
       title ="CDF for quality of service by load", caption = "Note: Heavy load is defined as bytesReceived in the top 5%,\n normal is below.")  + 
  theme(legend.position = "top", title = element_text(size = 9))

ggsave("final_plots/plot4_report.png", width = 8, height = 12, units = "cm")
# ggsave("final_plots/plot4_ppt.png", width = 8, height = 12, units = "cm")
```

Bandwidth by load
```{r}
lct <- Sys.getlocale("LC_TIME"); 
Sys.setlocale("LC_TIME", "C")

breakdown_load_ts %>% filter(time > "2018-09-05 00:00:00" & time < "2018-09-07 18:00:00" & variable != "bytesReceivedProxy") %>% 
  ggplot() + geom_area(aes(time, value, fill = load)) + 
  facet_grid(rows = vars(variable), scales = "free_y") +
  scale_fill_manual(values=c("red", "green")) +
    scale_x_datetime(date_breaks = "8 hours", date_labels = "%a\n%H") +
  labs(x = NULL, y = NULL,
       title = "Bandwidth and connections by load",
       subtitle = "(total bytes transferred and connections per 10 minutes)") +
  theme(legend.position = "top", title = element_text(size = 7))

ggsave("final_plots/plot5_report.png", width = 8, height = 16, units = "cm")

breakdown_load_ts %>% filter(time > "2018-09-05 00:00:00" & time < "2018-09-07 18:00:00" & variable %in% c("connections", "bytesReceived")) %>% 
  ggplot() + geom_area(aes(time, value, fill = load)) + 
  facet_grid(rows = vars(variable), scales = "free_y") +
  scale_fill_manual(values=c("red", "green")) +
  scale_x_datetime(date_breaks = "12 hours", date_labels = "%a\n%H") +
  labs(x = NULL, y = NULL,
       title = "Bandwidth and connections by load",
       subtitle = "(total bytes transferred and connections per 10 minutes)",
       caption = "Note: Heavy load is defined as bytesReceived in the top 5%,\n normal is below.") +
  theme(legend.position = "top", title = element_text(size = 7))

ggsave("final_plots/plot5_ppt.png", width = 10, height = 10, units = "cm")

Sys.setlocale("LC_TIME", lct)
```

CDF for CL_Retrans by load
```{r}
breakdown_load  %>% filter(variable == "CL_Retrans") %>% 
  ggplot() + geom_line(aes(value, CDF, colour = load), size = 2, alpha = 0.6) + scale_x_log10() + 
  scale_colour_manual(values=c("black", "red", "green")) +
  facet_wrap(vars(variable), nrow = 2, scales = "free_x") +
  labs(x = NULL, y = "cumulative distribution", 
       title ="CDF for re-transmissions\nbetween user and proxy by load", caption = "Note: Heavy load is defined as bytesReceived in the top 5%,\n normal is below.")  + 
  theme(legend.position = "top", title = element_text(size = 9))

ggsave("final_plots/plot6_report.png", width = 8, height = 8, units = "cm")
# ggsave("final_plots/plot6_ppt.png", width = 8, height = 12, units = "cm")
```

CDF connections per user in ten minutes

```{r}
p <- cdf_connections_per_user %>%
  #rename the feature from flows to connections
  mutate(variable = if_else(variable == "flows", "connections", variable),
         #reorder the factors
         variable = factor(variable, c("connections", "hosts", "domains"))) %>%
  ggplot() + geom_line(aes(value, CDF, colour = variable), size = 2) + scale_x_log10() + 
  scale_colour_manual(name = "measure",values = c("black", "orange", "red")) +
  labs(x = NULL, y = "cumulative distribution", 
       title = "CDF for connections per user\nin 10 minutes",
       caption = "Note: Counts on log scale.") + 
  theme(legend.position = "top")

p + theme(title = element_text(size = 8),
        legend.text = element_text(size = 7))
ggsave("final_plots/plot7_report.png", width = 8, height = 8, units = "cm")

p
ggsave("final_plots/plot7_ppt.png", width = 10, height = 10, units = "cm")
```

Number of connections started at once
```{r}
multi_count %>% mutate(number_fct = case_when(number < 7 ~ as.character(number),
                                              TRUE ~ "7-39")) %>% group_by(number_fct) %>% 
  summarise(observations = sum(observations)) %>% 
  ggplot() + geom_col(aes(x = number_fct, y = observations)) + scale_y_log10(labels = label_number()) + 
  coord_flip() + 
  labs(title = "Occurance of multiple connections\nstarted at the same time", y = "number of connections", x = "connections started at once",
       caption = "Note: Counts are on log scale.") + 
  theme(title = element_text(size = 9))

ggsave("final_plots/plot8_report.png", width = 8, height = 8, units = "cm")
# ggsave("final_plots/plot8_ppt.png", width = 8, height = 12, units = "cm")

```

CDF connections per host in 10 minutes
```{r}
cdf_connections_per_host %>% mutate(variable = if_else(variable == "flows", "connections", variable)) %>% 
  ggplot() + 
  geom_line(aes(value, CDF, colour = variable), size = 2) + scale_x_log10() + 
  scale_colour_manual(name = "measure",values = c("black", "orange")) +
  labs(x = NULL, y = "cumulative distribution", 
       title = "CDF for connections per host\nin 10 minutes",
       caption = "Note: Counts on log scale.") + 
  theme(legend.position = "top", title = element_text(size = 9))

ggsave("final_plots/plot9_report.png", width = 8, height = 8, units = "cm")
ggsave("final_plots/plot9_ppt.png", width = 10, height = 10, units = "cm")
```

CDF connections per user-host combination in 10 minutes
```{r}
cdf_connections_per_flow %>% 
  ggplot() + geom_line(aes(connections, CDF), size = 2) + scale_x_log10() + 
  labs(x= NULL, y = "cumulative distribution",
       title = "CDF for connections\nper user-host pair in 10 minutes",
       caption = "Note: Counts on log scale.") 

ggsave("final_plots/plot10_report.png", width = 8, height = 8, units = "cm")
ggsave("final_plots/plot10_ppt.png", width = 10, height = 10, units = "cm")
```
Direction of transfer ts

```{r}
lct <- Sys.getlocale("LC_TIME"); 
Sys.setlocale("LC_TIME", "C")

p3 <- breakdown_updown_ts  %>% filter(variable == "connections" & time < "2018-09-07 18:00:00" & time > "2018-09-05 00:00:00") %>% 
  ggplot() + geom_line(aes(time, value, colour = direction), size = 2) + 
  labs(x = NULL, y = "connections",
       title = paste("Time series for connections in 10 minutes\nby main direction")) +
  theme(legend.position = "top", title = element_text(size = 8))

p3 +  scale_x_datetime(date_breaks = "8 hours", date_labels = "%a\n%H")
ggsave("final_plots/plot11_report.png", width = 8, height = 8, units = "cm")

p3 + scale_x_datetime(date_breaks = "12 hours", date_labels = "%a\n%H")
ggsave("final_plots/plot11_ppt.png", width = 10, height = 10, units = "cm")

Sys.setlocale("LC_TIME", lct)
```

bandwidth

```{r}
lct <- Sys.getlocale("LC_TIME"); 
Sys.setlocale("LC_TIME", "C")


p4 <- breakdown_updown_bandwidth_ts %>% filter(time > "2018-09-05 00:00:00" & time < "2018-09-07 18:00:00" & variable == "bytesSent")%>% 
  ggplot() + geom_area(aes(time, value, fill = direction)) +
  scale_y_continuous(labels = label_bytes()) +
  labs(x = NULL, y = NULL,
       title = "Upload bandwidth by main direction of transfer",
       subtitle = "(data transferred in 10 minutes)") +
  theme(legend.position = "top", title = element_text(size = 8),
        legend.text = element_text(size = 7))

p4 + scale_x_datetime(date_breaks = "8 hours", date_labels = "%a\n%H")
ggsave("final_plots/plot12_report.png", width = 8, height = 8, units = "cm")

p4 +  scale_x_datetime(date_breaks = "12 hours", date_labels = "%a\n%H")
ggsave("final_plots/plot12_ppt.png", width = 10, height = 10, units = "cm")

Sys.setlocale("LC_TIME", lct)
```

UNADA-DGCA 1
```{r}
p <- ggplot(scores_by_flow) + geom_point(aes_string(x = "n_hosts",
                                             y = "n_con_user",
                                             color = "anomalous")) +
      theme_dark() + 
      scale_colour_manual(values = c("#FFFFFF", "#FF0040")) +
      labs(x = "count of distinct hosts in 10 minutes", y = "connections per user in 10 minutes") +
      theme(legend.position = "top")

p + theme(title = element_text(size = 8),
            legend.text = element_text(size = 7),
            axis.title.x = element_text(size = 7), axis.title.y = element_text(size = 7)) +
  labs(title = "Heavy connections analysed\nusing UNADA-DGCA")
ggsave("final_plots/plot13_report.png", width = 8, height = 8, units = "cm")

p
ggsave("final_plots/plot13_ppt.png", width = 10, height = 10, units = "cm")
```

UNADA-DGCA 2
```{r}
p2 <- ggplot(scores_by_flow) + geom_point(aes_string(x = "CL_RTT",
                                             y = "CL_Retrans",
                                             color = "anomalous")) +
      theme_dark() + 
      scale_colour_manual(values = c("#FFFFFF", "#FF0040")) +
      labs(x = "return-trip delays user-proxy", y = "re-transmissions user-proxy") +
      theme(legend.position = "top")

p2 + theme(title = element_text(size = 8),
          legend.text = element_text(size = 7),
          axis.title.x = element_text(size = 7), axis.title.y = element_text(size = 7)) +
  labs(title = "Heavy connections analysed\nusing UNADA-DGCA")
ggsave("final_plots/plot14_report.png", width = 8, height = 8, units = "cm")

p2
ggsave("final_plots/plot14_ppt.png", width = 10, height = 10, units = "cm")
```

Select 4 CDFs for ppt:
```{r}

cdfs_all %>% filter(variable %in% c("bytesSent", "CL_RTT", "CL_Retrans", "download_speed_OS_clean")) %>% 
  ggplot() + geom_line(aes(value, CDF)) + 
  scale_x_log10() + 
  facet_wrap(~variable, nrow = 2, ncol = 2, scales = "free_x") +
  labs(x = NULL, y = "cumulative distribution",
       title = "CDF for selected QoS measures",  caption = "Note: values on log scale.") 

ggsave("final_plots/plot_CDFs_ppt.png", width = 22, height = 14, units = "cm")

```

Traffic of most visited hosts
```{r}
top100_hosts_query <- "WITH TOTALS AS (SELECT COUNT(*) AS all_obs FROM logs.raw_view_clean),

HOSTS AS (SELECT hostFull, host2, COUNT(*) n_obs, ROW_NUMBER() OVER (ORDER BY COUNT(*) DESC) AS rank
    FROM logs.raw_view_clean 
      GROUP BY hostFull, host2
        ORDER BY n_obs DESC)
        
SELECT hostFull, host2, rank, n_obs, n_obs/all_obs AS share
  FROM HOSTS CROSS JOIN TOTALS
    LIMIT 100"

top100_hosts <- bq_project_query("case-study-2-nec-team", top100_hosts_query) %>% bq_table_download()

```

```{r}
top100_hosts %>% filter(rank < 21) %>% 
  ggplot() + geom_col(aes(x = rank, y = share, fill = host2)) + scale_y_continuous(labels = label_percent()) +
  guides(fill = "none") + 
  labs(title = "Traffic share of most-visited host IPs",
       subtitle = "(colored by domain)")

ggsave("final_plots/share_hosts_ppt.png", width = 10, height = 10, units = "cm")

```

