---
title: "Summary statistics detailed"
output: html_notebook
---

```{r include=FALSE}
library(tidyverse)
library(DBI)
library(bigrquery)
```

Let's retrieve CDFs for all variables of the cleaned dataset:

```{r}
con <- DBI::dbConnect(bigquery(), 
                      project = "case-study-2-nec-team",
                      dataset = "logs")

table <-  tbl(con, "raw_view")

vars <- table$ops$vars
numeric_vars <- vars[c(3:5, 10:27)]
```

```{r}
cdf_query <- paste("SELECT", 
                   paste0("approx_quantiles(", numeric_vars, ", 100) AS ", numeric_vars, collapse = ", "),
                   "FROM `case-study-2-nec-team.logs.raw_view`")

cdfs_all <- bq_project_query("case-study-2-nec-team", cdf_query) %>% 
  bq_table_download() %>% 
  # convert arrays into multiple rows
  unnest(., cols = names(.)) %>% 
  # add the comulative densities
  mutate(CDF = seq(0, 1, 0.01)) %>%
  # reshape to long format to use in charts
  pivot_longer(., names(.)[1:length(names(.))-1], names_to = "variable") %>% 
  group_by(variable) %>% arrange(CDF) %>% ungroup()
```
CDFs for all flows

```{r}
for (name in numeric_vars) {
p <- cdfs_all %>% filter(variable == name) %>% 
  ggplot() + geom_line(aes(value, CDF)) + labs(title = paste("CDF for", name)) 
print(p)
}
```

All CDFs are affected by extreme outliers.

Let's try again filtering out the maxima:

```{r}
for (name in numeric_vars) {
p <- cdfs_all %>% filter(variable == name & CDF < 1) %>% 
  ggplot() + geom_line(aes(value, CDF)) + labs(title = paste("CDF for", name, "until the 99%-ile")) 
print(p)
}
```

Statistics per user:
```{r}
aggregate_user <- paste("SELECT userID,", 
                        paste0("AVG(", numeric_vars, ") AS ", numeric_vars, collapse = ", "),
                   "FROM `case-study-2-nec-team.logs.raw_view` GROUP BY userid")

cdf_user_query <- paste("WITH agg_user AS (", aggregate_user, ") SELECT",
                        paste0("approx_quantiles(", numeric_vars, ", 100) AS ", numeric_vars, collapse = ", "),
                   "FROM agg_user")

cdfs_by_user <- bq_project_query("case-study-2-nec-team", cdf_user_query) %>% 
  bq_table_download() %>% 
  # convert arrays into multiple rows
  unnest(., cols = names(.)) %>% 
  # add the comulative densities
  mutate(CDF = seq(0, 1, 0.01)) %>%
  # reshape to long format to use in charts
  pivot_longer(., names(.)[1:length(names(.))-1], names_to = "variable") %>% 
  arrange(variable, CDF)

```

```{r}
for (name in numeric_vars) {
p <- cdfs_by_user %>% filter(variable == name) %>% 
  ggplot() + geom_line(aes(value, CDF)) + labs(title = paste("CDF for", name, "aggregated by user")) 
print(p)
}
```

```{r}
for (name in numeric_vars) {
p <- cdfs_by_user %>% filter(variable == name & CDF < 1) %>% 
  ggplot() + geom_line(aes(value, CDF)) + labs(title = paste("CDF for", name, "until 99%-ile, aggregated by user")) 
print(p)
}
```

Let's aggregate by host domain (host2)

```{r}
aggregate_host <- paste("SELECT host2,", 
                        paste0("AVG(", numeric_vars, ") AS ", numeric_vars, collapse = ", "),
                   "FROM `case-study-2-nec-team.logs.raw_view` GROUP BY host2")

cdf_host_query <- paste("WITH agg_host AS (", aggregate_user, ") SELECT",
                        paste0("approx_quantiles(", numeric_vars, ", 100) AS ", numeric_vars, collapse = ", "),
                   "FROM agg_host")

cdfs_by_host <- bq_project_query("case-study-2-nec-team", cdf_host_query) %>% 
  bq_table_download() %>% 
  # convert arrays into multiple rows
  unnest(., cols = names(.)) %>% 
  # add the comulative densities
  mutate(CDF = seq(0, 1, 0.01)) %>%
  # reshape to long format to use in charts
  pivot_longer(., names(.)[1:length(names(.))-1], names_to = "variable") %>% 
  arrange(variable, CDF)
```

```{r}
for (name in numeric_vars) {
p <- cdfs_by_host %>% filter(variable == name) %>% 
  ggplot() + geom_line(aes(value, CDF)) + labs(title = paste("CDF for", name, "aggregated by host")) 
print(p)
}
```

```{r}
for (name in numeric_vars) {
p <- cdfs_by_host %>% filter(variable == name & CDF < 1) %>% 
  ggplot() + geom_line(aes(value, CDF)) + labs(title = paste("CDF for", name, "until 99%-ile, aggregated by host domain")) 
print(p)
}
```

Let's have a look at developments over time (hours)

```{r}
time_query <- paste("SELECT DATETIME_TRUNC(Ts, HOUR) AS time, count(*) AS connections,", 
                        paste0("AVG(", numeric_vars, ") AS ", numeric_vars, collapse = ", "),
                   "FROM `case-study-2-nec-team.logs.raw_view` GROUP BY time")

time_series <- bq_project_query("case-study-2-nec-team", time_query) %>% 
  bq_table_download() %>% arrange(time) %>% filter(time >= "2018-09-05 02:00:00") %>% 
  pivot_longer(names(.)[2:length(names(.))], names_to = "variable")

```

```{r}
for (name in numeric_vars) {
p <- time_series %>% filter(variable == name | variable == "connections") %>% 
  ggplot() + geom_line(aes(time, value)) + facet_grid(rows = vars(variable), scales = "free_y") +
  labs(title = paste("time series for", name, "and connections")) 
print(p)
}
```

For some variables developments are not well visible due to the extreme event on 7 September. Let's filter this out

```{r}
for (name in numeric_vars) {
p <- time_series %>% filter((variable == name | variable == "connections") & time < "2018-09-07 18:00:00") %>% 
  ggplot() + geom_line(aes(time, value)) + facet_grid(rows = vars(variable), scales = "free_y") +
  labs(title = paste("time series for", name, "and connections")) 
print(p)
}
```

share of most important hosts:
```{r}
share_query <- "WITH TOTAL AS (SELECT count(*) AS flows FROM logs.raw_view),
  PER_HOST AS (SELECT host2, count(*) AS flows_per_host FROM logs.raw_view 
                  GROUP BY host2)
SELECT host2, flows_per_host / flows * 100 AS share 
  FROM PER_HOST CROSS JOIN TOTAL 
        ORDER BY flows_per_host DESC"

share_host <- bq_project_query("case-study-2-nec-team", share_query) %>% 
  bq_table_download() 
```

```{r}
availability_query <- paste0("SELECT ", 
                        paste0("COUNT(", numeric_vars, ") AS ", numeric_vars, collapse = ", "), " ",
                        "FROM `case-study-2-nec-team.logs.raw_view`")

availability <- bq_project_query("case-study-2-nec-team", availability_query) %>% bq_table_download() %>% 
  pivot_longer(., names(.), names_to = "variable", values_to = "count") 

availability %>% 
  #remove computed variables 
  # .[c(1:7, 10:11), ] %>% 
  mutate(variable = fct_reorder(variable, count)) %>% 
  ggplot() + geom_col(aes(x = variable, y = count)) + coord_flip()
```

